---
title: "Cristian C EDA"
format: html
---

```{r}
#load packages
library(tidyverse)
library(readr)
```

```{r}

# Download latest version
stadium_data <- read_csv("../../data/2022HitterStats.csv")

```

```{r}
# view basics of data
summary(stadium_data) #get summary of values of each variable

head(stadium_data) #view what the data looks like

```

##These is the codebook for stadium_data

team_name
ball_park
left_field - Distance from the home plate to 'the nearest fence or obstruction' in left field, measured in yards.
center_field - Distance from the home plate to 'the nearest fence or obstruction' in center field, measured in yards.
right_field - Distance from the home plate to 'the nearest fence or obstruction' in right field, measured in yards.
min_wall_height - Minimum wall height, measured in yards.
max_wall_height - Maximum wall height, measured in yards.
hr_park_effects - Total "park effects" on home run rate, with 100 being a baseline average.
extra_distance - Total extra distance that a particular ballpark adds to hits.
avg_temp - Average temperature at each ballpark, measured in Fahrenheit.
elevation - Elevation height above sea level, measured in feet.
roof - Percentage of all games that are played with roof cover on.
daytime - Percentage of all games that are played during the day.
