---
title: "Cristian C EDA"
format: html
---

```{r}
#load packages
library(baseballr)
library(tidyverse)
library(ggplot2)
library(readr)
library(dplyr)
library(ggridges)


parkfactor22 <- fg_park(2022)
parkfactor22

```

```{r}
#Stadium data
stadium_data <- read_csv("../../data/ballparks_stadium_data_2022.csv")

```


```{r}
# view basics of data
summary(stadium_data) #get summary of values of each variable
head(stadium_data) #view what the data looks like

stadium_data |>
  ggplot(aes(y = hr_park_effects,  x = team_name, fill = team_name)) +
  geom_col()

stadium_data |>
  ggplot(aes(x = extra_distance, y = hr_park_effects, color = ballpark)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    title = "Extra Distance vs HR Park Effects by Ballpark",
    x = "Extra Distance (yards)",
    y = "HR Park Effects (100 = league avg)"
  ) +
  theme_minimal()

```

##These is the codebook for stadium_data

team_name
ball_park
left_field - Distance from the home plate to 'the nearest fence or obstruction' in left field, measured in yards.
center_field - Distance from the home plate to 'the nearest fence or obstruction' in center field, measured in yards.
right_field - Distance from the home plate to 'the nearest fence or obstruction' in right field, measured in yards.
min_wall_height - Minimum wall height, measured in yards.
max_wall_height - Maximum wall height, measured in yards.
hr_park_effects - Total "park effects" on home run rate, with 100 being a baseline average.
extra_distance - Total extra distance that a particular ballpark adds to hits.
avg_temp - Average temperature at each ballpark, measured in Fahrenheit.
elevation - Elevation height above sea level, measured in feet.
roof - Percentage of all games that are played with roof cover on.
daytime - Percentage of all games that are played during the day.


**##However we later join in into joined_stadium_data with the 2022 hitter stats**
The code to do this are below


```{r}
`2022HitterStats` <- read_csv("../../data/2022HitterStats.csv")
head(`2022HitterStats`) 

```

```{r}
#player_id info
Player_team_info <- mlb_stats(stat_type = 'season', stat_group = 'hitting', season = 2022) |>
  select(player_id, team_name, league_name)
head(Player_team_info)

#Joined data set player_id with rest of player stats
Full_22_Hitter_stats <- `2022HitterStats` |>
  left_join(Player_team_info, join_by(player_id))
head(Full_22_Hitter_stats)

```


```{r}
#expected vs observed slg by hitter (colored by Park)
ggplot(Full_22_Hitter_stats, aes(x = xslg, y = slg_percent, color = team_name)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(x = "Expected SLG", y = "Observed SLG",
       title = "Observed vs. Expected SLG by Hitter (colored by Park)") +
  theme_classic()


```
```{r}
#Density graph of On base % by different teams
ggplot(Full_22_Hitter_stats, aes(x = on_base_percent, y = team_name, fill = team_name)) +
  geom_density_ridges(scale = 2, alpha = 0.7, size = 0.3) +
  labs(x = "On Base %",
       y = "Team",
       title = "Distribution of OBP by Home Park (Ridgeline)") +
  theme(legend.position = "none")

```

```{r}
#new dataset National_League_West_22 includes only Western teams
National_League_West_22 <- Full_22_Hitter_stats |>
  filter(team_name == "San Diego Padres" | team_name == "Los Angeles Dodgers" | team_name == "San Francisco Giants" | team_name == "Arizona Diamondbacks" | team_name == "Colorado Rockies")

ggplot(National_League_West_22, aes(x = xslg, y = slg_percent, color = team_name)) +
  geom_point()  +
    geom_smooth(method = lm, se = FALSE) +
  labs(x = "Expected SLG", y = "Observed SLG", color = "Team Name",
       title = "Observed vs. Expected SLG by Hitter (colored by Park) in National League West teams") +
  theme_classic()

```



```{r}
#Elevation column graph in different stadiums
stadium_data |>
  ggplot(aes(x = ballpark, y = elevation, fill = ballpark)) +
  geom_col()

```

```{r}
#Joining stadium data
joined_stadium_data <- stadium_data |>
  left_join(Full_22_Hitter_stats, join_by(team_name))
head(joined_stadium_data, 20)

joined_stadium_data |>
  group_by(team_name) |>
  summarise(mean(home_run))

joined_stadium_data |>
  filter(team_name == "San Diego Padres" | team_name == "Los Angeles Dodgers" | team_name == "San Francisco Giants" | team_name == "Arizona Diamondbacks" | team_name == "Colorado Rockies") |>
  ggplot(aes(x = xslg, y = slg_percent, color = team_name)) +
  geom_point()

#Filter for teams with greater than certain ammount of observations
# joined_stadium_data |>
#   count(team_name) |>
#   filter(n > 4)


```
```{r}
#Some preliminary graphs

# 1) Actual vs Expected Batting Average
ggplot(joined_stadium_data, aes(x = xba, y = batting_avg, color = league_name)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.7, size = 2.5) +
  labs(
    title = "Actual vs. Expected Batting Average",
    x = "Expected BA (xba)",
    y = "Observed BA (batting_avg)"
  ) +
  theme_minimal()

# 2) Actual vs Expected Slugging
ggplot(joined_stadium_data, aes(x = xslg, y = slg_percent, color = league_name)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.7, size = 2.5) +
  labs(
    title = "Actual vs. Expected Slugging %",
    x = "Expected SLG (xslg)",
    y = "Observed SLG (slg_percent)"
  ) +
  theme_minimal()

# 3) Park Effects vs. Hard-Hit % 
ggplot(joined_stadium_data, aes(x = hr_park_effects, y = hard_hit_percent, color = ballpark)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "HR Park Effects vs. Hard-Hit %",
    x = "HR Park Effects (100 = avg)",
    y = "Hard-Hit %"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# 4) Extra Distance vs Home Runs
ggplot(joined_stadium_data, aes(x = extra_distance, y = home_run, color = ballpark)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed") +
  labs(
    title = "Extra Distance vs. Home Runs",
    x = "Extra Distance (yards)",
    y = "Home Runs"
  ) +
  theme_minimal() +
  theme(legend.position = "none")


```
