---
title: "Proposal"
---
```{r}
#load need packages
library(tidyverse)
```

### Team Name: Heyyy Batter Batter

## Team Members:

  - Cristian Cano
  - Wandi Dlamini
  - Izzy Flood
  - Zach Marshak
  
## Description

Main Question: How do Park factors (how teams/players are predicted to perform at different stadiums) explain deviations between predicted stats values, such as home runs, versus actual performance in the MLB

We will be using the Baseballr package, which pulls data from various databases. In our project, we will mainly be using data from Baseball Savant, StatCast, and FanGraphs. 

## Inspiration/Reasons

Izzy has been a baseball player and fan his entire life, he explained what park factors are to the rest of our group and we all found it very interesting.


## Responsibility Plan

```{r}
# Create a data frame with group members and their duties
project_team <- tibble::tribble(
  ~Member_Name, ~Responsibility,
  "Cristian",   "Case Study, Visualization work",
  "Wandi",   "Figuring out the BaseballR package, data vis work",
  "Izzy", "Baseball expert, report writer",
  "Zach",     "Proposal scribe, cleaning and wrangling")

# Display the table
print(project_team)
```
=======


